Joint maximum-likelihood of phylogenies and ancestral states is not consistent

### 2018-05-15 joint_inf.tex
Adding a clarifying sentence, and clearing out :+1: comments.

@@

\begin{figure}
\centering
% obtained by running
% python joint_inf_plot.py --analytic --topology --delta .001 --plot-name figures/topology-inconsistency.svg
\includegraphics[width=\textwidth]{topology-inconsistency-inkscape}
\caption{
%EM I'm sure you're getting to this, but we need to change around wording to compensate for the use of probabilities vs. fidelities now.
%das: You just mean for the figure captions, right? I changed it in the figures and a few places in the text, but maybe you meant we should change around the whole text to be in terms of branch lengths? Should we change the InvFels/Fels tree figure, too?
    An analytically-derived region of topological inconsistency in terms of probabilities of a character change along a branch for ``perfect'' data generated on the InvFels topology (Fig.~\ref{fig:farris-fels-top}) with $p_{w^*} = p_{y^*}$ (in terms of fidelities, $w^*=y^*$).
    Due to the looseness of the upper and lower bounds, the parameters in the white region do not necessarily indicate consistency, though all parameters in the shaded region result in an inconsistency.
}
\label{fig:inconsistency-farris}
\end{figure}

@@

%EM Let's set a time in the first half of this week for you to walk me through the Goldman paper and this paragraph. Thanks!
The case of joint inference of a phylogenetic likelihood was discussed in \citet{Goldman1990-dk}.
In it, Goldman provides a worked example in which estimating a topology with fixed branch lengths is equivalent to parsimony and thus not guaranteed to be consistent, though he does not discuss the inconsistency of joint inference in general.
We show cases where the estimates of the topology will be inconsistent if we are estimating both topology and branch lengths, and, moreover, even if we know the correct topology, we show cases where branch lengths will also be inconsistent.
Finally, just prior to his conclusion, he discusses when parsimony will give the same answer as maximum likelihood, concluding that the question is ill-posed since parsimony estimates different parameters than maximum likelihood, i.e., it assumes equal branch lengths.
The joint inference procedure outlined here estimates the same parameters as classical maximum likelihood---albeit implicitly estimating ancestral states as well---rendering the question well-posed.
Finally, we are able to provide much more detail on how large branch lengths must be for general joint inference---estimating topology and branch lengths together---to fail to be consistent.

>>>>

Although we use fidelities exclusively for our theoretical development, we have made our figures in terms of probabilities of substitution $p_{\theta} = (1-\theta)/2$ as they are easier to interpret.

@@

\begin{figure}
\centering
% obtained by running
% python joint_inf_plot.py --analytic --topology --delta .001 --plot-name figures/topology-inconsistency.svg
\includegraphics[width=\textwidth]{topology-inconsistency-inkscape}
\caption{
    An analytically-derived region of topological inconsistency in terms of probabilities of a character change along a branch for ``perfect'' data generated on the InvFels topology (Fig.~\ref{fig:farris-fels-top}) with $p_{w^*} = p_{y^*}$ (in terms of fidelities, $w^*=y^*$).
    Due to the looseness of the upper and lower bounds, the parameters in the white region do not necessarily indicate consistency, though all parameters in the shaded region result in an inconsistency.
}
\label{fig:inconsistency-farris}
\end{figure}

@@

The case of joint inference of a phylogenetic likelihood was discussed in \citet{Goldman1990-dk}.
In it, Goldman provides a worked example in which estimating a topology with fixed branch lengths is equivalent to parsimony and thus not guaranteed to be consistent, though he does not discuss the inconsistency of joint inference in general.
We show cases where the estimates of the topology will be inconsistent if we are estimating both topology and branch lengths, and, moreover, even if we know the correct topology, we show cases where branch lengths will also be inconsistent.
Finally, just prior to his conclusion, he discusses when parsimony will give the same answer as maximum likelihood, concluding that the question is ill-posed since parsimony estimates different parameters than maximum likelihood, i.e., it assumes equal branch lengths.
The joint inference procedure outlined here estimates the same parameters as classical maximum likelihood---albeit implicitly estimating ancestral states as well---rendering the question well-posed.
Finally, we are able to provide much more detail on how large branch lengths must be for general joint inference---estimating topology and branch lengths together---to fail to be consistent.
